---
title: 'Shift of choice towards EV: A comparative analysis'
author: "Group_4"
date: "2023-12-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```

#Loading dataset
```{r}
Vehicle <- read_excel("~/Library/CloudStorage/OneDrive-NortheasternUniversity/Project R/Group 4/Vehicle_dataset .xlsx")
Path <- "~/Library/CloudStorage/OneDrive-NortheasternUniversity/Project R/Group 4/Vehicle_dataset .xlsx"
```
#Dataset sorting
```{r}
sheet_names <- excel_sheets(Path)
total_data <- read_excel(Path, sheet = sheet_names[8])
total_data <- as.data.frame(total_data)
colnames(total_data) <- c("Year", "EV sales", "Alternative Vehicle sales", "Conventional Vehicle sales","EV+Alt", "Electric Sales Difference", "Alternative Sales Difference", "Conventional Sales Difference")
```
#Sorted Data
```{r}
print(total_data)
```

#Data from 2016-2019 Pre-covid
```{r}
population_data <- total_data[total_data$Year<2020,]
population_data$`EV sales` <- as.numeric(gsub(",","",population_data$`EV sales`))
population_data$`Alternative Vehicle sales` <- as.numeric(gsub(",","",population_data$`Alternative Vehicle sales`))
population_data$`Conventional Vehicle sales` <- as.numeric(gsub(",","",population_data$`Conventional Vehicle sales`))
print(population_data)
```

#Data from 2020-2022 Post-covid
```{r}
sample_data <- total_data[total_data$Year>2019,]
sample_data$`EV sales` <- as.numeric((gsub(",","", sample_data$`EV sales`)))
sample_data$`Alternative Vehicle sales` <- as.numeric((gsub(",","", sample_data$`Alternative Vehicle sales`)))
sample_data$`Conventional Vehicle sales` <- as.numeric((gsub(",","", sample_data$`Conventional Vehicle sales`)))
print(sample_data)
```

#Mean of Electric sales and Conventional sales
```{r}
sample_mean <- mean(total_data$`Electric Sales Difference`)
print(sample_mean)
population_mean <- mean(total_data$`Conventional Sales Difference`)
print(population_mean)
```

#Pre and post data 
```{r}
pre_data <- total_data[total_data$Year < 2020, ]
post_data <- total_data[total_data$Year >2019, ]
```
#To check if there is significance difference between mean of EV Sales and mean of conventional vehicle sales for the all years
```{r}
t.test(total_data$`Electric Sales Difference`, mu = population_mean, alternative = "two.sided", conf.level = 0.95)
```
# Checking if the EV sales mean is significantly different from the conventional sales between 2016 and 2019
```{r}
t.test(pre_data$`Electric Sales Difference`, pre_data$`Conventional Sales Difference`, 
       mu = mean(pre_data$`Conventional Sales Difference`), alternative = "two.sided",
       conf.level = 0.95)
```

# Checking if the EV sales mean is significantly different from conventional sales between 2020 and 2022
```{r}
t.test(post_data$`Electric Sales Difference`, post_data$`Conventional Sales Difference`,
       mu = mean(post_data$`Conventional Sales Difference`), alternative = "two.sided",
       conf.level = 0.95)
```
# Inference: Between 2016 and 2019 there is a significant difference between EV sales and conventional vehicle sales,however in the second analysis of sales between 2020 and 2021, there seems to be insufficient proof to accurately reject the null hypothesis. Hence to conclude the sales mean after COVID hit is not significantly different thereby indicating how buyer preference for EV is climbing with a decline in conventional vehicle preferences.

# Create a new data frame with relevant columns
```{r}
plot_data <- data.frame(Year = total_data$Year,
                        Electric_Difference = total_data$`Electric Sales Difference`,
                        Conventional_Difference = total_data$`Conventional Sales Difference`)
```

# Plotting for Electric Sales
```{r}
ggplot(plot_data, aes(x = Year, y = Electric_Difference)) +
  geom_point(color = "blue", size = 3, shape = 16) +
  geom_line(color = "blue", linetype = "solid", size = 1) +
  labs(title = "Electric Sales Differences Over Time",
       x = "Year",
       y = "Electric Sales Difference") +
  theme_minimal()
```

# Plotting for Conventional Sales
```{r}
ggplot(plot_data, aes(x = Year, y = Conventional_Difference)) +
  geom_point(color = "red", size = 3, shape = 17) +
  geom_line(color = "red", linetype = "dashed", size = 1) +
  labs(title = "Conventional Sales Differences Over Time",
       x = "Year",
       y = "Conventional Sales Difference") +
  theme_minimal()
```

#Plots showing Trend line of Conventional sales and Electric sales
```{r}
plot_data <- data.frame(Year = total_data$Year,
                        Electric_Difference = total_data$`Electric Sales Difference`,
                        Conventional_Difference = total_data$`Conventional Sales Difference`)
ggplot(plot_data, aes(x = Year)) +
  geom_point(aes(y = Electric_Difference), color = "blue", size = 3, shape = 16) +
  geom_line(aes(y = Electric_Difference), color = "blue", linetype = "solid", size = 1) +
  geom_point(aes(y = Conventional_Difference), color = "red", size = 3, shape = 17) +
  geom_line(aes(y = Conventional_Difference), color = "red", linetype = "dashed", size = 1) +
  geom_text(aes(x = max(Year), y = max(Electric_Difference), label = "Electric Sales"),
            color = "blue", hjust = 1.2, vjust = 0.5) +
  geom_text(aes(x = max(Year), y = max(Conventional_Difference), label = "Conventional Sales"),
            color = "red", hjust = 1.2, vjust = 0.5) +
  labs(title = "Electric and Conventional Sales Differences Over Time",
       x = "Year",
       y = "Sales Difference") +
  scale_shape_manual(values = c(16, 17)) +
  scale_linetype_manual(values = c("solid", "dashed")) +
  scale_color_manual(values = c("blue", "red")) +
  theme_minimal()
```

# Pre COVID Pie chart
```{r}
pie_data_pre <- data.frame(X = c("EV sales", "Conventional sales"),
                       Y = c(sum(pre_data$`Electric Sales Difference`),
                             sum(pre_data$`Conventional Sales Difference`)),
                       Z = c((sum(pre_data$`Electric Sales Difference`)/sum(pre_data$`Electric Sales Difference`, pre_data$`Conventional Sales Difference`))*100,
                             (sum(pre_data$`Conventional Sales Difference`)/sum(pre_data$`Electric Sales Difference`, pre_data$`Conventional Sales Difference`))*100))
pie_data_pre$Z <- round(pie_data_pre$Z, 0)
ggplot(pie_data_pre, aes(x = "", y = Y, fill = X)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  geom_text(aes(label = paste0(Z, "%")), position = position_stack(vjust = 0.5))+
  ggtitle("Proportion of Sales pre COVID") +
  theme_void()                       
```

# Post COVID Pie chart
```{r}
pie_data <- data.frame(X = c("EV sales", "Conventional sales"),
                       Y = c(sum(post_data$`Electric Sales Difference`),
                             sum(post_data$`Conventional Sales Difference`)),
                       Z = c((sum(post_data$`Electric Sales Difference`)/sum(post_data$`Electric Sales Difference`, post_data$`Conventional Sales Difference`))*100,
                             (sum(post_data$`Conventional Sales Difference`)/sum(post_data$`Electric Sales Difference`, post_data$`Conventional Sales Difference`))*100))
pie_data$Z <- round(pie_data$Z, 0)
ggplot(pie_data, aes(x = "", y = Y, fill = X)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  geom_text(aes(label = paste0(Z, "%")), position = position_stack(vjust = 0.5))+
  ggtitle("Proportion of Sales post-covid") +
  theme_void()                       
```
## Plotting trend lines for 5 states
```{r}
state_ev_data <- read_excel("~/Library/CloudStorage/OneDrive-NortheasternUniversity/Project R/Group 4/State_wise_EV.xlsx")
state_ev_data <- as.data.frame(state_ev_data)
trend_line_data <- data.frame(X = state_ev_data$Year,
                              S1 = state_ev_data$MA,
                              S2 = state_ev_data$NY,
                              S3 = state_ev_data$GE,
                              S4 = state_ev_data$NJ,
                              S5 = state_ev_data$FL)
ggplot(trend_line_data, aes(x = X))+
  geom_point(aes(y = S1), color ="blue", size = 3, shape = 15)+
  geom_line(aes(y = S1), color = "blue")+
  geom_text(aes(x = max(X), y = max(S1), label = "MA"), color = "blue", hjust = 0.5, vjust = 2)+
  geom_point(aes(y = S2), color ="red", size = 3, shape = 14)+
  geom_line(aes(y = S2), color = "red")+
  geom_text(aes(x = max(X), y = max(S2), label = "NY"), color = "red", hjust = 1.2, vjust = 0.5)+
  geom_point(aes(y = S3), color ="green", size = 3, shape = 16)+
  geom_line(aes(y = S3), color = "green")+
  geom_text(aes(x = max(X), y = max(S3), label = "GE"), color = "green", hjust = 1.2, vjust = -0.5)+
  geom_point(aes(y = S4), color ="brown", size = 3, shape = 17)+
  geom_line(aes(y = S4), color = "brown")+
  geom_text(aes(x = max(X), y = max(S4), label = "NJ"), color = "brown", hjust = 1.2, vjust = 0.5)+
  geom_point(aes(y = S5), color ="orange", size = 3, shape = 18)+
  geom_line(aes(y = S5), color = "orange")+
  geom_text(aes(x = max(X), y = max(S5), label = "FL"), color = "orange", hjust = 1.2, vjust = 0.5)+
  labs(title = "Trend of EV sales of New England States",
       x = "Year",
       y = "Sales")+
  theme_minimal()
```

# Creating trend lines to verify conventional vehicle distribution
```{r}
state_nev_data <- read_excel("~/Library/CloudStorage/OneDrive-NortheasternUniversity/Project R/Group 4/State_wise_nonEV.xlsx")
state_nev_data <- as.data.frame(state_nev_data)
trend_line_ndata <- data.frame(X = state_nev_data$Year,
                               S1 = state_nev_data$MA,
                               S2 = state_nev_data$NY,
                               S3 = state_nev_data$GE,
                               S4 = state_nev_data$NJ,
                               S5 = state_nev_data$FL)
ggplot(trend_line_ndata, aes(x = X))+
  geom_point(aes(y = S1), color ="blue", size = 3, shape = 15)+
  geom_line(aes(y = S1), color = "blue")+
  geom_text(aes(x = max(X), y = S1[7], label = "MA"), color = "blue", hjust = 0.5, vjust = 2)+
  geom_point(aes(y = S2), color ="red", size = 3, shape = 14)+
  geom_line(aes(y = S2), color = "red")+
  geom_text(aes(x = max(X), y = S2[7], label = "NY"), color = "red", hjust = 1.2, vjust = 0.5)+
  geom_point(aes(y = S3), color ="green", size = 3, shape = 16)+
  geom_line(aes(y = S3), color = "green")+
  geom_text(aes(x = max(X), y = S3[7], label = "GE"), color = "green", hjust = 0.25, vjust = -0.75)+
  geom_point(aes(y = S4), color ="brown", size = 3, shape = 17)+
  geom_line(aes(y = S4), color = "brown")+
  geom_text(aes(x = max(X), y = S4[7], label = "NJ"), color = "brown", hjust = 0.5, vjust = -1)+
  geom_point(aes(y = S5), color ="orange", size = 3, shape = 18)+
  geom_line(aes(y = S5), color = "orange")+
  geom_text(aes(x = max(X), y = S5[7], label = "FL"), color = "orange", hjust = 1.2, vjust = 0.5)+
  labs(title = "Trend of Non EV sales of New England States",
       x = "Year",
       y = "Sales")+
  theme_minimal()
```